dispRity manual

### 2020-04-22 inst/gitbook/01_glossary.Rmd
Updated gitbook until section 3

In this manual | In `dispRity` | E.g. in palaeobiology | E.g. in ecology
---------------|---------------|------------------|-------------
the multidimensional space | a `matrix` object ($n\times k$) | a morphospace | a function-space
elements | rows ($n$) | taxa | field experiments
dimensions | columns ($k$) | morphological characters | communities' compositions
subsets | a `matrix` ($m \times k$, with $m \leq n$) | time series | experimental treatments
disparity | a `function` | sum of variances | ellipsoid volume

>>>>

In this manual | In `dispRity` | E.g. in palaeobiology | E.g. in ecology
---------------|---------------|------------------|-------------
the multidimensional space | a `matrix` object ($n\times d$) | a morphospace | a function-space
elements | rows ($n$) | taxa | field experiments
dimensions | columns ($d$) | morphological characters | communities' compositions
subsets | a `matrix` ($m \times d$, with $m \leq n$) | time series | experimental treatments
disparity | a `function` | sum of variances | ellipsoid volume

### 2020-04-22 inst/gitbook/02_getting-started.Rmd
Updated gitbook until section 3

@@

Classically, when a high number of variables is used, disparity is calculated from ordinated matrices.
These can be any type of ordinations (PCO, PCA, PCoA, MDS, etc.) as long as elements are the rows (taxa, countries, field experiments) and the dimensions are the columns.

@@

> This functionality is not available on the CRAN version of the package (`r version_CRAN`) due to CRAN compatibility issues see [to CRAN or not to CRAN?](#noCRAN) for more information and on how to install the GitHub (more complete) version of the package (`r version_release`).

@@

`dispRity` package can easily take data from `Claddis` using the `Claddis.ordination` function.
For this, simply input a matrix in the `Claddis` format to the function and it will automatically calculate and ordinate the distances among taxa:

@@

Of course you can use your own data as detailed in the [previous chapter](#What-sort-of-data-does-dispRity-work-with).

@@

 * An matrix with rows as elements and columns as dimensions (always!)
 * A list of group members: this list should be a list of numeric vectors or names corresponding to the row names in the matrix. For example `list("a" = c(1,2), "b" = c(3,4))` will create a group _a_ containing elements 1 and 2 from the matrix and a group _b_ containing elements 3 and 4. Note that elements can be present in multiple groups at once.
 * Your favourite disparity metric (here the sum of variances)

>>>>

Since version `1.4` it is also possible to include a `"list"` containing matrices.
These matrices need to have the same dimensions and rownames but can contain different values.
This is especially useful for modelling uncertainty (see [here](#multi.input) for more details).

@@

Classically, when a high number of variables is used, disparity is calculated from ordinated matrices.
These can be any type of ordinations (PCO, PCA, PCoA, MDS, etc.) as long as elements are the rows (taxa, countries, field experiments) and the dimensions are the columns.
However, note that this is not required from any of the functions in this package.
You can also use distance matrices or any other matrix type that suits your question and your analysis!

@@

@@

`dispRity` package can also easily take data from the `Claddis` package using the `Claddis.ordination` function.
For this, simply input a matrix in the `Claddis` format to the function and it will automatically calculate and ordinate the distances among taxa:

@@

Of course you can use your own data as detailed in the [previous section](#What-sort-of-data-does-dispRity-work-with).

@@

 * An matrix with rows as elements and columns as dimensions (always!)
 * A list of group members: this list should be a list of numeric vectors or names corresponding to the row names in the matrix. For example `list("A" = c(1,2), "B" = c(3,4))` will create a group _A_ containing elements 1 and 2 from the matrix and a group _B_ containing elements 3 and 4. Note that elements can be present in multiple groups at once.
 * Your favourite disparity metric (here the sum of variances)

### 2020-04-22 inst/gitbook/index.Rmd
Updated gitbook until section 3

```{r, eval = TRUE, echo = FALSE}
## master version
version_raw <- RCurl::getURL("https://raw.githubusercontent.com/TGuillerme/dispRity/master/DESCRIPTION")
version_master <- strsplit(strsplit(version_raw, split = "Version: ")[[1]][2], split = "\nDate:")[[1]][1]

## release version
version_raw <- RCurl::getURL("https://raw.githubusercontent.com/TGuillerme/dispRity/release/DESCRIPTION")
version_release <- strsplit(strsplit(version_raw, split = "Version: ")[[1]][2], split = "\nDate:")[[1]][1]

## CRAN version
version_raw <- RCurl::getURL("https://cran.r-project.org/web/packages/dispRity/index.html")
version_CRAN <- strsplit(strsplit(version_raw, split = "\n<td>Version:</td>\n<td>")[[1]][2], split = "</td>\n</tr>\n<tr>\n<td>Depends:")[[1]][1]
```

@@

This is a package for measuring disparity in `R`.
It allows users to summarise matrices as representations as multidimensional spaces into a single value or distribution describing a specific aspect of this multidimensional space (the disparity).
Multidimensional spaces can be ordinated matrices from MDS, PCA, PCO, PCoA but the package is *not* restricted to any type of matrices!
This manual is based on the version `r version_release`.

>>>>

```{r, eval = TRUE, echo = FALSE}
## master version
version_raw <- RCurl::getURL("https://raw.githubusercontent.com/TGuillerme/dispRity/master/DESCRIPTION")
version_master <- strsplit(strsplit(version_raw, split = "Version: ")[[1]][2], split = "\nDate:")[[1]][1]
version_master <- 1.4

## release version
version_raw <- RCurl::getURL("https://raw.githubusercontent.com/TGuillerme/dispRity/release/DESCRIPTION")
version_release <- strsplit(strsplit(version_raw, split = "Version: ")[[1]][2], split = "\nDate:")[[1]][1]
version_release <- 1.4

## CRAN version
version_raw <- RCurl::getURL("https://cran.r-project.org/web/packages/dispRity/index.html")
version_CRAN <- strsplit(strsplit(version_raw, split = "\n<td>Version:</td>\n<td>")[[1]][2], split = "</td>\n</tr>\n<tr>\n<td>Depends:")[[1]][1]
version_CRAN <- 1.4
```

@@

This is a package for measuring disparity (aka multidimensional space occupancy) in `R`.
It allows users to summarise matrices as representations as multidimensional spaces into a single value or distribution describing a specific aspect of this multidimensional space (the disparity).
Multidimensional spaces can be ordinated matrices from MDS, PCA, PCO, PCoA but the package is *not* restricted to any type of matrices!
This manual is based on the version `r version_release`.
